Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-06-12 articles/formalism/sections/real_space_formalism.tex
Editing through visualizing kernels

%
\begin{figure}[!t]
\centering
\includegraphics[width=0.5\columnwidth]{radiation_convolution.pdf}
\caption{\revisit{The local cartesian coordinates $(\hat{x},\hat{y})$ are drawn on the red circle(sphere), representative of the coordinate dependence of the Stokes parameters. The two sets of dotted lines drawn at representative points denote great circles, one which passes through the central point labelled `e' and the other chosen such that the two have locally orthogonal tangent vectors $(\hat{r},\hat{\phi})$. The angle $\alpha_{qe}$ defines a rotation operator that aligns the local $\hat{x}$ with $\hat{r}$.   \textit{Convolution kernel:} The resultant scalar field at `e' can be evaluated by summing over the contribution from all the Stokes parameters on the red circle. This convolution is performed with kernels which are defined in term of the Euler angle $\gamma_{eq}$. \textit{Radiation kernel:} It is also possible to compute the contribution from the Stokes parameter at `q' to all the point on the blue circle and this is a function of the Euler angle $\alpha_{qe}$. On the flat sky $\gamma_{eq}=-\alpha_{qe}$ and consequently there is no difference between the radiation and convolution kernels.}}
\label{fig:planar_euler_angles}
\end{figure}
%

@@

%--------------------------------------------------------
\subsection{Evaluating Stokes $Q$/$U$ from scalar $E$/$B$}\label{sec:eb2qu}
The real space operator which translates $E$ \& $B$ fields to Stokes parameters $Q$ \& $U$ can be derived using a similar procedure. Expressed in the matrix-vector notation, inverse operator is given by the following equation:
%
\begin{subequations}
\beqry
\bar{P} &=& \bar{T}^{-1} {{}_2\mathcal{Y}}\, \tilde T {{}_0\mathcal{Y}^{\ddagger}}\bar{S} = \frac{1}{2} \bar{T}^{\dagger} {{}_2\mathcal{Y}} \,\tilde T {{}_0\mathcal{Y}^{\ddagger}}\bar{S}\,,  \\
&=&  \bar O^{-1} \bar{S}\,.
\eeqry
\end{subequations}
%
%The inverse operator $\bar{O}^{-1}$ can be derived by realizing that $\mm^{-1}$ is given by the following expression,
%%
%\beq
%\mm^{-1} = \sum_{\ell m}  {{}_{2}}Y_{\ell m}(\hat n_q) {}_{0}Y^*_{\ell m}(\hat n_e) \, 
%\eeq
%%
The inverse operator expressed in terms of the function $\mm$ given in \eq{eq:qu2eb_gen_kernel} is given by the following equation:
%
\beq
{\bar O}^{-1}=- \Delta \Omega\bmat \mathcal{M}_{r} & -\mathcal{M}_{i} \\  \mathcal{M}_{i}  & \mathcal{M}_{r} \emat=- \Delta \Omega{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{qe}) & -\sin(2\alpha_{qe})\\  \sin(2 \alpha_{qe})  & \cos(2 \alpha_{qe}) \emat \,,
\eeq
%
where all the symbols have the same meaning as discussed in \sec{sec:qu2eb}. Note that the kernel in the above equation differs from the one in \eq{eq:op_qu2eb_rad} by a change in sign on the off-diagonals of the block matrix. When expressed in terms of the same set of Euler angles used to define the operator $\bar{O}$, it can be shown that the different forms of the real space operator are given by the following equations:
%
\beqry
    {}_{+2}X(\hat{n}_q) &&=  \sum_{e=1}^{N_{\rm pix}} [E+iB](\hat{n}_{e})\   \mathcal{M}^*_{B}[\hat{n}_e] \hspace{0.8cm}\textrm{\emph{Radiation kernel}}\,,\label{eq:eb2qu_radiation} \\
    {}_{+2}X(\hat{n}_q) &&= \Bigg\lbrace \mathcal{M}^*_{G} \star [E+iB] \Bigg\rbrace(\hat{n}_q) \hspace{0.8cm}\textrm{\emph {Convolution kernel}}\,, \label{eq:eb2qu_convolution}
\eeqry
%
where all the symbols have the same meaning as defined in \sec{sec:qu2eb}. Note that the conjugated forms of the Green's function and the effective beam for the operator $\bar{O}$ have their roles reversed for the inverse operator $\bar{O}^{-1}$.
%--------------------------------------------------------

@@

The real and imaginary part of the kernel $\mathcal{M}_G$ are identical irrespective of changes in the galactic latitude and longitude of the central pixel. These functions are not distorted when a part of the domain overlaps with the poles, as can be seen in the first two rows of \fig{fig:vis_kernel}. These observations are consistent with the Green's function interpretation of the kernel $\mathcal{M}_{G}$ which yield coordinate independent scalars E/B.  From \eq{eq:qu2eb_radiation_concise} it is clear that $\mathcal{M}_G$ can be interpreted in the following ways:
%
\beq
\bmat E= \textrm{Re}(\mathcal{M}_G) \\ B =\textrm{Im}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=\delta(\hat{n} - \hat{n}_q)\\ U=0 \emat \hspace{5 mm};\hspace{5 mm}
\bmat E= -\textrm{Im}(\mathcal{M}_G) \\ B = + \textrm{Re}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=0 \\ U=\delta(\hat{n} - \hat{n}_q) \emat\,,
\eeq
%
$\mathcal{M}^*_B$ is the Greens function of the operator that transforms E/B to Stokes parameters Q/U. The real and imaginary parts of $\mathcal{M}^*_B$ are identical irrespective of the galactic longitude, however they vary as a function of galactic latitude. This latitude dependent shape is necessary as the coordinate dependence of the Stokes parameters derived from the scalar E/B is built in via this Greens function. From \eq{eq:eb2qu_radiation} it is clear that $\mathcal{M}^*_B$ can be interpreted in the following manner:
%
\beq
\bmat Q= \textrm{Re}(\mathcal{M}^*_B) \\ U = \textrm{Im}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=\delta(\hat{n} - \hat{n}_e)\\ B=0 \emat \hspace{5 mm};\hspace{5 mm}
\bmat Q= -\textrm{Im}(\mathcal{M}^*_B) \\ U = \textrm{Re}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=0 \\ B=\delta(\hat{n} - \hat{n}_e) \emat\,.
\eeq
%
%Both these facts can be associated with the fact that this function does not depend on the Euler angle $\gamma$. 
%\comment{Add discussion here on the inverse greens function and why it changes as a function of position.}

>>>>

%
\begin{figure}[!t]
\centering
\includegraphics[width=0.5\columnwidth]{radiation_convolution.pdf}
\caption{\revisit{The local cartesian coordinates $(\hat{x},\hat{y})$ are drawn on the red circle(sphere), representative of the coordinate dependence of the Stokes parameters. The two sets of dotted lines drawn at representative points denote great circles, one which passes through the central point labelled `e' and the other chosen such that the two have locally orthogonal tangent vectors $(\hat{r},\hat{\phi})$. The angle $\alpha_{qe}$ defines a rotation operator that aligns the local $\hat{x}$ with $\hat{r}$.   \textit{Convolution kernel:} The resultant scalar field at `$e$' can be evaluated by summing over the contribution from all the Stokes parameters on the red circle. This convolution is performed with kernels which are defined in term of the Euler angle $\gamma_{eq}$. \textit{Radiation kernel:} It is also possible to compute the contribution from the Stokes parameter at `$q$' to all the point on the blue circle and this is a function of the Euler angle $\alpha_{qe}$. On the flat sky $\gamma_{eq}=-\alpha_{qe}$ and consequently there is no difference between the radiation and convolution kernels.}}
\label{fig:planar_euler_angles}
\end{figure}
%

@@

%--------------------------------------------------------
\subsection{Evaluating Stokes $Q$/$U$ from scalar $E$/$B$}\label{sec:eb2qu}
The real space operator which translates $E$/$B$ fields to Stokes parameters $Q$/$U$ can be derived using a similar procedure. Expressed in the matrix-vector notation, inverse operator is given by the following equation:
%
\begin{subequations}
\beqry
\bar{P} &=& \bar{T}^{-1} {{}_2\mathcal{Y}}\, \tilde T {{}_0\mathcal{Y}^{\ddagger}}\bar{S} = \frac{1}{2} \bar{T}^{\dagger} {{}_2\mathcal{Y}} \,\tilde T {{}_0\mathcal{Y}^{\ddagger}}\bar{S}\,,  \\
&=&  \bar O^{-1} \bar{S}\,.
\eeqry
\end{subequations}
%
%The inverse operator $\bar{O}^{-1}$ can be derived by realizing that $\mm^{-1}$ is given by the following expression,
%%
%\beq
%\mm^{-1} = \sum_{\ell m}  {{}_{2}}Y_{\ell m}(\hat n_q) {}_{0}Y^*_{\ell m}(\hat n_e) \, 
%\eeq
%%
The inverse operator expressed in terms of the function $\mm$ given in \eq{eq:qu2eb_gen_kernel} is given by the following equation:
%
\beq
{\bar O}^{-1}=- \Delta \Omega\bmat \mathcal{M}_{r} & -\mathcal{M}_{i} \\  \mathcal{M}_{i}  & \mathcal{M}_{r} \emat=- \Delta \Omega{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{qe}) & -\sin(2\alpha_{qe})\\  \sin(2 \alpha_{qe})  & \cos(2 \alpha_{qe}) \emat \,,
\eeq
%
where all the symbols have the same meaning as discussed in \sec{sec:qu2eb}. Note that the kernel in the above equation differs from the one in \eq{eq:op_qu2eb_rad} by a change in sign on the off-diagonals of the block matrix. When expressed in terms of the same set of Euler angles used to define the operator $\bar{O}$, it can be shown that the different forms of the real space operator are given by the following equations:
%
\beqry
    {}_{+2}X(\hat{n}_q) &&=  \sum_{e=1}^{N_{\rm pix}} [E+iB](\hat{n}_{e})\   \mathcal{M}^*_{B}[\hat{n}_e] \hspace{0.8cm}\textrm{\emph{Radiation kernel}}\,,\label{eq:eb2qu_radiation} \\
    {}_{+2}X(\hat{n}_q) &&= \Bigg\lbrace \mathcal{M}^*_{G} \star [E+iB] \Bigg\rbrace(\hat{n}_q) \hspace{0.8cm}\textrm{\emph {Convolution kernel}}\,, \label{eq:eb2qu_convolution}
\eeqry
%
where all the symbols have the same meaning as defined in \sec{sec:qu2eb}. Note that the conjugated forms of the radiation kernel (Green's function) and the convolution kernel (effective beam) for the operator $\bar{O}$ have their roles reversed for the inverse operator $\bar{O}^{-1}$.
%--------------------------------------------------------

@@

The kernel $\mathcal{M}_G$ is the Green's function of the operator that transforms Stoke parameters to coordinate independent $E/B$.  The coordinate independence implies that the real and imaginary parts of the kernel do not vary with changes in the galactic latitude and longitude of the central pixel. In particular these functions are not distorted when a part of the domain overlaps with the poles, as can be seen in the first two rows of \fig{fig:vis_kernel}. From \eq{eq:qu2eb_radiation_concise}, we can see the $E/B$ patterns that $\delta$-functions in the Stokes parameters produce:
%
\beq
\bmat E= \textrm{Re}(\mathcal{M}_G) \\ B =\textrm{Im}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=\delta(\hat{n} - \hat{n}_q)\\ U=0 \emat ;\qquad
\bmat E= -\textrm{Im}(\mathcal{M}_G) \\ B = + \textrm{Re}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=0 \\ U=\delta(\hat{n} - \hat{n}_q) \emat.
\eeq
%
On the other hand, $\mathcal{M}^*_B$ is the Green's function of the operator that transforms E/B to coordinate-dependent Stokes parameters Q/U. The kernel $\mathcal{M}^*_B$ does not change with the central longitude, but varies as a function of galactic latitude. This latitude-dependent shape carries the  coordinate dependence of the Stokes parameters. From \eq{eq:eb2qu_radiation}, we $\mathcal{M}^*_B$ arises from $\delta$-functions in $E/B$:
%
\beq
\bmat Q= \textrm{Re}(\mathcal{M}^*_B) \\ U = \textrm{Im}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=\delta(\hat{n} - \hat{n}_e)\\ B=0 \emat;\qquad
\bmat Q= -\textrm{Im}(\mathcal{M}^*_B) \\ U = \textrm{Re}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=0 \\ B=\delta(\hat{n} - \hat{n}_e) \emat\,.
\eeq
%
%Both these facts can be associated with the fact that this function does not depend on the Euler angle $\gamma$. 
%\comment{Add discussion here on the inverse greens function and why it changes as a function of position.}
